---
title: "LAWS-HiC Example"
output: html_document
date: "2024-11-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Input requirement

The required input of LAWS-HiC are:

* results from a standard peak caller (e.g., FitHiC2): only bin pair coordinators and p-value are needed for LAWS-HiC

* domain list, in this example domain list is topologically associated domain (TAD) list

Let's first look at the input format of the standard peak caller results:

```{r}
library(data.table)
example_data <- fread("./example_data.tsv")
example_domain_list <- fread("./example_domain_list.tsv")

example_data[1:10,]
```

The four columns of the standard peak caller results should be:

* chr: chromosome number

* fragmentMid1, fragmentMid2: coordinators of bin pairs (middle position of the fragment in this example), 

* p_value: p-value from the standard peak caller

The format for the domain list:

```{r}
example_domain_list[1:10,]
```

The four columns are chromosome number, start and end position of the domain.
The four columns of the domain list should be:

* chr: chromosome number 

* x1: start position of the domain

* x2: end position of the domain



## LAWS-HiC

The required parameters of LAWS-HiC function:

* input: data of standard peak caller results with format previously shown

* domain_input: data of domain list with format previously shown

* chr: chromosome number

* resolution: resolution of the original in kb

```{r}
source("./laws_hic.R")
example_output <- laws_hic(input=example_data, 
                           domain_input=example_domain_list, 
                           chr=21, resolution=10)
```

This function will output the tuned tau and h for each domain.

```{r}
example_output[1:10,]
```


The columns in the output data are: 

* chr:chromosome number

* fragmentMid1, fragmentMid2: coordinators of bin pairs (same as input)

* p_value: p-value from the standard peak caller

* i,j: identifiers of bin pairs

* p_laws: LAWS-HiC adjusted p-value

* pi: LAWS-HiC estimated spatial level

